A modelling strategy to improve cacao quality and productivity

### 2021-11-14 paperModel/paperModel_2.tex
Corrections

%=================================================================
\documentclass[gene,journal,article,submit,moreauthors,pdftex]{Definitions/mdpi} 
\usepackage{natbib}
\usepackage[colorinlistoftodos]{todonotes}
\newcounter{mycomment}
\newcommand{\mycomment}[2][]{%
% initials of the author (optional) + note in the margin
\refstepcounter{mycomment}%
{%
\setstretch{0.7}% spacing
\todo[color={red!100!green!33},size=\small]{%
\textbf{Comment [\uppercase{#1}\themycomment]:}~#2}%
}}
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

@@

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

@@

% Authors, for the paper (add full first names)
\Author{Angela Romero V $^{1,\ddagger}$*\orcidA{}, Anyela V. Camargo R. $^{1, \ddagger}$ and Adriana M. Gallego$^{2}$}

@@

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Romero, A.; Gallego, A.; Camargo, A.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{$^{1}$ \quad The John Bingham Laboratory, NIAB, 93 Lawrence Weaver Road, Cambridge CB3 0LE, UK 1\\$^{2}$ \quad Grupo BIOS, Centro de Bioinformática y Biología Computacional de Colombia - BIOS, Manizales, Caldas, Colombia, South America. }%; e-mail@e-mail.com}

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{Crop modelling can support agronomical decisions of crop production under a range of scenarios improving competitiveness. Cocoa production systems in Latin America has a high importance over social and economic development, facing  the fight against hunger and poverty. Although Colombian cocoa has the potential to be in the high value markets for fine flavour, it is still not widely produced as the lack of adoptions of technologies by the traditional farmers. They empirically harvest after 5 or 6 months after flowering date. However, cocoa fruits development can be considered as the result of a number of physiological and morphological processes that can be described by mathematical relationships even under uncontrolled environments. Thus, we parametrized the SIMPLE crop model \citep{Zao2019simple} to predict the best time for harvest cocoa fruits in Colombia. The results showed an RRMSE of 7.2\% for the yield prediction, while the simulated harvest date varied between +/- 2 to 20 days depending on the temperature variations of the year between regions. This crop model application contributed to understand and predict the phenology of cacao fruits the varieties ICS95 y CCN51, which is key to produce high quality cacao beans.  The aim of this study was developed a practical tool for farmers on cocoa fields to predict the best moment to harvest, using easily available data for crop modelling such as flowering date and weather variables (solar radiation, rain, and temperature). }

@@

\section{Introduction}
Cocoa ( \textit{Theobroma cacao }L. ) is an important worldwide perennial tropical crop endemic to the South American rainforests \citep{zuidema2005, motamayor2002, argout2011, Rodriguez2019}. Cacao plant is a member of the Malvaceae (formerly Sterculiaceae)  botanical family such as  cotton \textit{ Gossypium hirstium} \citep{Nix2017cotton}. Coton has been is modeled in SIMPLE model \citep{Zao2019simple}. Cocoa is grown for its fruits, known as cacao pods \citep{ Niemenak2010, suarez2021}. Only the 5\% of the world cocoa yield is destined for Fine-cocoa production due to the low productivity through  the traditional crop management \citep{argout2011}.  In Colombia, cocoa  is  traditionally  consumed  as  a  beverage. It is one of the crops promoted by the Colombian government in the social and agricultural development  programs aimed at favouring peace in post-conflict regions \citep{Rodriguez2019, Abbott2019}. This crop is grown by approximately 52.000 families \citep{Gutierrez2020} and 98\% of production being carried out by small and medium-sized producers \citep{Garcia2014, Escobar2020}. Colombia registered an increase of 3.750 tons in production in 2020 compared to the previous year \citep{lamos2020}.

Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}, it is still not widely produced as the lack of adoptions of technologies by the traditional farmers. They empirically harvest at 5 or 6 months after flowering date, hence   a mix of quality of cocoa beans can be fermented. This produce heterogeneous characteristics between each fermentation batch diminishing the quality of cocoa final product \citep{Escobar2021}. To identify the best moment to harvest is important to consider physiological responses affected by climate variables such as rain, solar radiation and wind.  Thus,  for cocoa in Colombia, physiological simulation models may be valuable to identify the best moment to harvest cocoa considering variable weather conditions, soil types and cultivar specifications.

@@

In this paper, we present a physiological parametrization of SIMPLE crop model for cocoa to predict the best harvest time  and yield production. We used the SIMPLE crop model \citep{Zao2019simple} for three reasons: 1: That it is very comprehensively described in the original paper. 2: That the code was available in R for initial trials and 3: That it had already been successfully fitted to other perennial crops in south America. Overall, the model simulates crop development, growth and yield, and predict the maturation day when the fruit is ready to harvest. It includes 13 parameters ( daily weather data, irrigation, and soil and key dates) to specify a crop type, with four of these for cultivar characteristics easily available from farmer. Thus, this could be used as  tool for small farmers with the aim to improve the quality of cocoa to become more competitive in Fine-cocoa market.

\section{Materials and Methods}
\subsection{Floral Phenology of Cocoa }
Normally the phenological stages in cocoa are  divided in two main phases: vegetative and reproductive. In the SIMPLE model we simulate the reproductive phase (fig. \ref{fig:pheno}, a) described by the floral phenology from the date of inflorescence emergence (BBCH scale 5) (fig. \ref{fig:pheno},b ) to predict the date of ripening of fruit and seed (BBCH 8) (fig. \ref{fig:pheno},c ) \citep{Niemenak2010} . In the Andean region the reproductive phase is cyclically fulfilled during two annual cycles passing by the following phases: inflorescence emergence, flowering, pollination, fruit development and harvest. Therefore, for modelling parametrization the crop cycle of cocoa as perennial plant does not start at the plantation date such as annual crops systems. Instead  the start point of the cocoa crop cycle to model is the inflorescence emergence date (fig. \ref{fig:pheno},b). Consequently , the growth period of the fruit can varied from  110 to 150  daa (days after anthesis) \citep{lopez2018} when cacao fruits reaches the physiological maturity, but it can be harvested at 170 days daa \citep{Niemenak2010} for quality purposes.

Cocoa is a cauliflorous plant, which means that flowers grow on the trunk and branches. In Colombia cocoa trees usually produce flowers throughout the year. Cocoa trees produces with up to 10000 flowers per tree each year, which the 50 \%  do not develop into ripe fruits according to Fedecacao reports. The flower takes 30 days passing by 12 micro-stages from meristem development (stages 1to 6) to the fully developed flower (Stages 7 to 12) \citep{swanson2005} when it is ready to be pollinated. The opening of flowers or anthesis  occurs over a 12-hour period during the night and it is synchronised between the groups of mature flowers \citep{Niemenak2010}. However, the live of a flower can last approximately 1 day after the opening falling form the trunk if it is unfertilised \citep{cheesman1927, Niemenak2010}.

Subsequentially, after anthesis the fruit growths by approximately 150 days until the maturation, mucilage. Therefore, the complete maturation process of the fruit, from the pollination to fully mature fruit, takes 160- 210 days \citep{berry1994}. The accumulation of lipids, storage proteins and anthocyanin starts about 85 days after pollination when fruits have an active metabolism and seeds   moisture content decreases up to 30\% \citep{Lehrian1980, Niemenak2010}. During this phase the quality of cocoa seeds is defined.

@@

\subsection{Test site and yield production }
Cacao is cultivated in 30 states out of 32 from the total Colombian territory with about 147,000 ha \citep{Meza2021}. Cocoa fields studied were located in Saravena (Arauca), Rionegro (Santander), Cali (Valle del Cauca), Apartado (Antioquia) and Manizalez (Caldas) (Fig.\ref{fig:yield}, b). We considered 112 parcels data from Fedecacao reports. Each parcel data contained age of crop, density of planting, yield, number of fruits harvested, flowering date, harvest date.   Acording to personal comunication from framers the biggest flowering occurs in September and January to harvest in March and July. However, the data reported showed that this flowering and pod productions are not constant for all the regions.  Cali and Caldas presented the lowest production. However, the pod harvest in Caldas increased in May and from October to December, and decreased from January to March.  Meanwhile, Arauca and Apartado reported the highest yield in the months January, July, November and December. Santander has picks of production in March, May and September (Fig.\ref{fig:yield}, a).

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.3]{images/map.png}\\
	\caption{\footnotesize {Cocoa production characterization of cocoa for five locations\\}}
	\label{fig:yield}
\end{figure}
\newpage

\subsection{Weather conditions}
\begin{figure}[h!]
	\centering
	\includegraphics[scale=0.4]{images/clima.png}
	\caption{\footnotesize {Colombian weather conditions. \\}}	
	\label{fig:temp}
	{\footnotesize (a), Available photosynthetic solar radiation (PAR). (b), Daily average temperature  (c), Monthly average of precipitation (bars) and relative humidity (dotted line) from 2018 to 2021 }
\end{figure}
\newpage

Solar radiation (PAR), temperature, precipitation ant relative humidity was studied from 2018 to 2020 (Fig.\ref{fig:temp}). Santander and Caldas had the biggest variability and the maximum of solar radiation values over 20 MJ m$^{2}$day$^{-1}$. In contrast , Cali and Arauca presented the lowest values of PAR below 5 MJ m$^{2}$day$^{-1}$. (Fig.\ref{fig:temp}, a).  Even though, Cali and Santander had contrasting PAR conditions, they regions presented  have similar temperature during 2020 (Fig.\ref{fig:temp}, b). The temperature ranges from 16 to 28 $^\circ$C and it is relatively constant for each region. However, Arauca presented the biggest variability with hotter months during the first half of the year 2019 and 2020.  Apartado was found as the hottest region studied with 26 $^\circ$C and Caldas as the coldest site with 18 $^\circ$C.  Precipitation in Colombia is presented in two seasons per year from February to April and from October to November, while the relative humidity remain constant over 80\% (Fig.\ref{fig:temp},c). In general, the coldest regions tested (Caldas and Santander) had the maximun values of solar radiation available for photosynthesis.

\subsection{Thermal time for pod harvest date identification}
It is required to model crop growth  considering the cocoa base temperature (T$_{b}$), at which the plant development stops \citep{Ritchie1991}. This characterization was conducted for the five regions counted 180 days (6 months) after flowering, as farmers used to harvest by calendar days. Characterizing the thermal time the models predicts the maturation day to harvest the cocoa can vary depending on temperature variations. Thermal time (degree days) was calculated  from the flowering date to the pod harvest  using the ecuation \ref{equ:tem}. Where tt is the cumulative sum of the daily temperature (T$_{i}$) and T$_{b}$ for cocoa is 10$^\circ$C according to \cite{lahive2019}.

@@

Calibration of crop models are conducted typically for particular cultivars and require site specific inputs of weather \citep{Crout20142}. The procedure for the SIMPLE model \citep{Zao2019simple} calibration  was a sequential process of modifying physiological variables specific for cocoa in the inputs files and  adding the appropriate files of weather for each region tested. In the SIMPLE model has seven input files where new cocoa crop data should be provided. Files in the list below can be edited to define the features of the new cultivars or experiments. Cocoa crop has not been simulated with SIMPLE model, hence consider previous cocoa studies, values such as leaf area index (LAI) \citep{Agele2016} and Harvest index (HI) \citep{Quintana2015} where modified. Radiation Use Efficiency (RUE) was calibrate according to the perennial crops  (banana and cotton) that had been calibrated previously in SIMPLE model \citep{Zao2019simple} with a RUE of 0.8 and 0.85 respectively. As cocoa trees are under shadow the RUE was lower with values between 0.7 and 0.5 g MJ$^{\mathbf{-1}}$ m$^{2}$ (table \ref{tab:reparam}).
Once the physiological parameters were calibrated to the  simulated yields for cocoa were reasonably close to the observed yield. 23 flowering dates from 12-July-2019 to 23-June-2020 were introduced in the treatment file, running the program and saving the results.

\begin{enumerate}
	\item Input/Simulation Management.csv
	\item Input/Species parameter.csv
	\item Input/Cultivar.csv
	\item Input/Treatment.csv	
	\item Input/Soil.csv
	\item Observation/Obsdummy crop Exp name.csv	
	\item Weather/dummy weather.WTH
\end{enumerate}

\subsection{Inputs and parameters}
Input variables required to run SIMPLE model for cocoa include the flowering date and daily weather of solar radiation (SRAD), maximum and minimum temperature (TMAX, TMIN) and rain. Weather data as csv file was  downloaded from the POWER Data Access Viewer \citep{nasapower} from January 01 2018 to December 31 2020 for four locations in Colombia (Cali, Rionegro -Santander, Apartado - Antioquia, Saravena Arauca). The csv file had to be transformed to .WHT file using R 1.4 version \citep{Rstudio2020}.

There were three parameters varied by region (table \ref{tab:reparam}) thermal time required for pod harvest after the flowering date (Tsum) , the Radiation use efficiency (RUE) and yield observed on field. Physiological parameters in table \ref{tab:Treaparam} are  common  for all the regions studied. These parameter were calibrate for cultivars ICS95 and CCN51 considering a range of time of 200 days (DAP) from flowering date to harvest day, even thought farmers collect the pod at 180 DAP. Heat and water stress parameters were not considered.

@@

\subsection{Evaluation of model performance}

The SIMPLE model performance was evaluated by comparing simulated values cocoa yield with those reported by Fedecacao from cocoa plantations, using the statistical indice of relative root mean square error (RRMSE) (Equ. \ref{equ:RMSE}) \citep{Zao2019simple, Bai2020}.

@@

\section{Results}

\subsection{Weather effects over flowering time }

Figure \ref{fig:heat} shows the pearson correlation to study the weather data of the flowering time over the months of flowering  (monthF), month of harvest (monthH) and their final yield (fruit\_kg) . The results showed that thermal time with T$_{b}$ (ttb), temperature minimum and maximum (TMIN and TMAX) and Dew Frost Point at 2 meters (T2MDEW) are positive related. The wind (WS2M) is clear correlated with ttb. However, less clear correlations were found of monthF  with T2MDEW, relative humidity (RH), WS2M and rain.

\begin{figure}[h!]
	\centering
	\includegraphics[scale=0.35]{images/heatm.png}\\
	\caption{\footnotesize {Pearson correlation average weather variables and flowering dates for all regions. \\}}
	\label{fig:heat} 
\end{figure}
\newpage

\subsection{Thermal time }
Thermal time characterization  was made considering 180 days after flowering (daf) for each region. The boxplot in the figure \ref{fig:ttbox} shows the data distribution where boxes indicate the range of the central 50\% of the total data per region, the central line in the box is marking the median value and lines draw out from each box mean the range of the remaining data. Therefore, this boxplot shows differences between locations as was expected following the tendencies of the temperature  (fig. \ref{fig:temp},c). Apartado and Arauca had the highest temperatures hence, the highest thermal time values with 2909 and 2764 days$^\circ$C  respectively. Caldas had the lowest values with 1173 days$^\circ$C. Meanwhile, Cali and Santander presented similar thermal time around 2000 $^\circ$C (table \ref{tab:reparam}). The accumulated temperature during the pod development (fig.\ref{fig:ttbox}) depends on the region where cocoa is cultivated and the variety planted. Thermal time values are also proportional to the yield reported on field (table \ref{tab:reparam}).\\

\begin{figure}[h!]
	\centering
	\includegraphics[scale=0.3]{images/ttbbox.png}\\
	\caption{\footnotesize {Cocoa yield and thermal time characterization at 180 days after flowering.\\}}
	\label{fig:ttbox}
\end{figure}

\subsection{Model validation}
Therefore, biomass production of the aerial part of the plant was simulated per region (fig. \ref{fig:m1}, a) as biomass-rate which is the daily biomass growth rate.  As for cocoa simulation we did not consider water and heat stress variables, in this research biomass simulation is affected by the solar radiation, RUE, daily temperature, atmospheric CO$_{2}$ concentration (ppm) and the fraction of solar radiation intercepted by a tree of cocoa during the fruit development (fSolar) (fig. \ref{fig:m1}, b). Due to the lack of this kind of data from fields, the comparison was not possible with observed data of biomass production through the crop cycle. However, the biomass simulation also responded proportionally with the yield production. Biomass of cocoa fruits simulated include all the parts that compound  the pod when they are fresh. Therefore, these this information may be of interest to farmers to predict pod fresh mass.

>>>>

%=================================================================
\documentclass[gene,journal,article,submit,moreauthors,pdftex]{Definitions/mdpi} 
\usepackage{natbib}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[dvipsnames]{xcolor}
\newcounter{mycomment}
\newcommand{\mycomment}[2][]{%
% initials of the author (optional) + note in the margin
\refstepcounter{mycomment}%
{%
\setstretch{0.7}% spacing
\todo[color={red!100!green!33},size=\small]{%
\textbf{Comment [\uppercase{#1}\themycomment]:}~#2}%
}}
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

@@

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, characterisation, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

@@

% Authors, for the paper (add full first names)
\Author{Angela P. Romero V. $^{1,\ddagger}$*\orcidA{}, Anyela V. Camargo R.$^{1, \ddagger}$, Oscar D. Ramirez$^{3}$, Paula A. Arenas V$^{3}$ and Adriana M. Gallego$^{2}$}

@@

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Romero, A.P.V;  Camargo, A.; Arenas P.; Ramirez O. D.; Gallego, A.;}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{$^{1}$ \quad The John Bingham Laboratory, NIAB, 93 Lawrence Weaver Road, Cambridge CB3 0LE, UK 1\\$^{2}$ \quad Grupo BIOS, Centro de Bioinformática y Biología Computacional de Colombia - BIOS, Manizales, Caldas, Colombia, South America.\\ $^{3}$ \quad Federación Nacional de Cafeteros FEDECACAO, Colombia, South America. }%; e-mail@e-mail.com}

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{Cacao production systems in Latin America {\color{red}have } a high importance over social and economic development, facing  the fight against hunger and poverty. Although Colombian cacao has the potential to be in the high value markets for fine flavour, {\color{red} the lack of expert support as well as the use of traditional, and oftentimes suboptimal, technologies makes cocoa production negligibly. For example, traditionally cacao harvest takes place at 5 or 6 months after flowering, other environmental parameters that have more association with pod maturation speed are not taken into account.  Cocoa fruits development can be considered as the result of a number of physiological and morphological processes that can be described by mathematical relationships even under uncontrolled environments. In this context, crop models are useful tools to simulate and predict crop development over time and under multiple environmental conditions. Since, harvesting at the right time can yield high quality cacao, we parametrised a crop model to predict the best time for harvest cocoa fruits in Colombia. Our aim was to develop a practical tool that supports cacao farmers in the production of high quality cacao.} When comparing simulated and observed data, our results showed an RRMSE of 7.2\% for the yield prediction, while the simulated harvest date varied between +/- 2 to 20 days depending on the temperature variations of the year between regions. {\color{red} This crop model contributed to understand and predict the phenology of cacao fruits for two key cacao varieties ICS95 y CCN51.}
}

@@

\section{Introduction}

Cocoa ( \textit{Theobroma cacao }L. ) is an important worldwide perennial tropical crop endemic to the South American rainforests \citep{zuidema2005, motamayor2002, argout2011, Rodriguez2019}. Cacao plant is a member of the Malvaceae (formerly Sterculiaceae)  botanical family such as  cotton \textit{ Gossypium hirstium} \citep{Nix2017cotton}. Coton has been is modeled in SIMPLE model \citep{Zao2019simple}. Cocoa is grown for its fruits, known as cacao pods \citep{ Niemenak2010, suarez2021}. Only the 5\% of the world cocoa yield is destined for Fine-cocoa production due to the low productivity through  the traditional crop management \citep{argout2011}.  In Colombia, cocoa  is  traditionally  consumed  as  a  beverage. It is one of the crops promoted by the Colombian government in the social and agricultural development  programs aimed at favouring peace in post-conflict regions \citep{Rodriguez2019, Abbott2019}. This crop is grown by approximately 52.000 families \citep{Gutierrez2020} and 98\% of production being carried out by small and medium-sized producers \citep{Garcia2014, Escobar2020}. Colombia registered an increase of 3.750 tons in production in 2020 compared to the previous year \citep{lamos2020}.

Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}, {\color{red} the lack of expert support as well as the use of traditional, and often times suboptimal, technologies makes cocoa production negligibly. For example, the farmer practice of empirically harvesting from 5 to 6 months or {\color{olive} 180 days } after flowering date (DAF), produced a mix of quality of cocoa beans that are then be fermented. This practice produces heterogeneous characteristics between each fermentation batch which potentially diminish the quality of the end cocoa product} \citep{Escobar2021}. To identify the best moment to harvest is important to consider physiological responses affected by climate variables such as rain, solar radiation and wind.  Thus,  for cocoa in Colombia, physiological simulation models may be valuable to identify the best moment to harvest cocoa considering variable weather conditions, soil types and cultivar specifications.

@@

In this paper, we present a physiological parametrisation of SIMPLE crop model for cocoa to predict best harvest time and overall yield. {\color{red} We parameterised the SIMPLE crop model \citep{Zao2019simple} because it had already been successfully fitted to other tree crops in south America. The model simulates crop development, growth and yield, and predicts the maturation day when the fruit is likely to be ready for harvest. Our aim was to develop a practical tool that supports cacao farmers in the production of high quality cacao. Thus, making Colombian cacao more competitive in the Fine-cocoa market.}

\section{Materials and Methods}

\subsection{Floral Phenology of Cocoa}

Usually, the phenological stages of a cacao tree are divided in two main phases: vegetative and reproductive. Defining the reproductive phase (fig. \ref{fig:pheno}, a) as described by the floral phenology from the date of inflorescence emergence (BBCH scale 5) (fig. \ref{fig:pheno},b ) to the date of ripening of fruit and seed (BBCH 8) (fig. \ref{fig:pheno},c ) \citep{Niemenak2010}. The reproductive phase in the Andean region of Colombia is cyclically fulfilled during two annual cycles which goes through the following phases: inflorescence emergence, flowering, pollination, fruit development and harvest. Therefore, for modelling parameterisation the crop cycle of cocoa as perennial plant does not start at the plantation date such as annual crops systems. Instead, the start point of the cocoa crop cycle is the inflorescence emergence date (fig. \ref{fig:pheno},b). Consequently , the growth period of the fruit can vary from  110 to 150  daa (days after anthesis) \citep{lopez2018} when cacao fruits reach the physiological maturity, but it can be harvested at 170 days daa \citep{Niemenak2010} for quality purposes.

{\color{red} Cocoa is a cauliflorous plant, which means that flowers grow on the trunk and branches. Cocoa trees usually produce up to 10000 flowers per tree each year, 50 \%  of them do not develop into ripe fruits (Personal communication with Fedecacao). Flower development takes approximately 30 days across 12 micro-stages, from} meristem development (stages 1 to 6) to the fully developed flower (Stages 7 to 12) \citep{swanson2005, lopez2018} when it is ready to be pollinated. The opening of flowers or anthesis  occurs over a 12-hour period during the night and it is synchronised between the groups of mature flowers \citep{Niemenak2010}. However, the live of a flower can last approximately 1 day after the opening falling form the trunk if it is unfertilised \citep{cheesman1927, Niemenak2010}. {\color{red} Subsequentially, after anthesis, fruits growth for approximately 150 days until the maturation, mucilage}. Therefore, the complete maturation process of the fruit, from the pollination to fully mature fruit, takes 160- 210 days \citep{berry1994}. The accumulation of lipids, storage proteins and anthocyanin start about 85 days after pollination when fruits have an active metabolism and seeds moisture content decreases up to 30\% \citep{Lehrian1980, Niemenak2010}. During this phase the quality of cocoa seeds is defined.

@@

\subsection{Inputs and Data Acquisition }

{\color{olive} Input variables required to run the cocoa model included the flowering date and daily weather of solar radiation (SRAD), maximum and minimum temperature (TMAX, TMIN) and rain. Therefore, weather data as csv file was  downloaded from the POWER Data Access Viewer \citep{nasapower} from January 01 of 2018 to December 31 of 2020 for the five locations in Colombia (Fig.\ref{fig:yield}, b). The csv file had to be transformed to .WHT file using R 1.4 version \citep{Rstudio2020}.

Field data for this study was provide by Fedecacao  (National Cacao Producers Federation). Reports contained information of 23 flowering dates from 12-July-2019 to 23-June-2020. Each of these dates had  their corresponding date of harvest after at exactly 180 DAF,  the age of the trees, plant density (trees ha$^{-1}$), yield  (dry beans kg ha$^{-1}$) and number of fruits harvested per hectare. Thus, 23 flowering dates for five farms gave us 115 samples in total.}

\subsection{Test Site and Yield Production }

{\color{olive} Five farms in different regions were tested : Saravena (Arauca), Rionegro (Santander), Cali (Valle del Cauca), Apartado (Antioquia) and Manizalez (Caldas). These were considered because had the data available of production in the reports of  Fedecacao. Yield depend on the successfully development of flowers to form ripe pods. } {\color{red}  According to personal communication from farmers, the highest flowering season occurs in September and January suggesting that harvest occurs in March and July. However, the data collected from the farms suggested that flowering and pod production were not constant for all the locations. For example, pod harvest in Caldas farm increased in May and from October to December, and decreased from January to March.  Meanwhile, the Arauca and Apartado farms reported the highest yield in the months January, July, November and December. Santander farm had picks of production in March, May and September (Fig.\ref{fig:yield}, a).} {\color{olive} However,  cacao is cultivated in 30 regions of Colombian which accounts for approximately about 147,000 ha \citep{Meza2021}. Therefore, the national production of cocoa beans in 2020 was  63.416 ton according  to  information  published by Fedecacao (\citep{Fede2021}). The highest yield was reposted by Santander region with 26,315,  following in descending order by Antioquia with 5.974 tons, Arauca with 5.082 tons, Caldas with 1.343 tons and Valle del Cauca with 339 tons.}

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.3]{images/map.png}\\
	\caption{\footnotesize {Cocoa production  of five farms in different regions.\\}}
	\label{fig:yield}
	{\footnotesize (a) Production per month (2019 - 2020) of five farms. (b) Map showing the regions where farms are located.}
\end{figure}
\newpage

\subsection{Thermal Time for Pod Harvest Date Identification}

{\color{olive} The cumulative sum of daily temperature from a reference day 0 is defined as \emph{Thermal time} and  its units of measurement are in days degrees (days $^\circ$C). That starter point of 0 days $^\circ$C generally is the planting date \citep{Ritchie1991} but for cocoa crop simulation model it is the flowering date.  Thermal time of a cultivated plant may consider the base temperature (T$_{b}$), which is the minimum temperature required by cocoa plant to grow. T$_{b}$ an vary between cultivars \citep{Slafer1995, Daymond2008}. For cocoa the vegetative growth T$_{b}$ has been reported between 18.6 and 20.8 $^\circ$C \citep{Daymond2008}. Nevertheless, the pod growth  has a lower T$_{b}$ which range between 9 and 12.9 $^\circ$C \citep{Daymond2008, lahive2019}. We calculated the cocoa thermal time with a pod growth T$_{b}$ of 10 $^\circ$C because it is the absolute minimum temperature for cocoa growing in South America reported by \cite{Erneholm1948}, in \citep{lahive2019}.

The thermal time required for the crop model was characterised for each location starting from flowering date (0 days $^\circ$C) to harvest date (180 after flowering = 6 months), as farmers used to harvest by calendar days. }  Thus, the cocoa model predicts the maturation day to harvest pods. It can vary depending on temperature variations.  Thermal time was calculated using the equation \ref{equ:tem}. Where tt is the cumulative sum of the daily temperature (T$_{i}$) and T$_{b}$ for cocoa is 10$^\circ$C.

@@

Calibration of crop models are conducted typically for particular cultivars and require site specific inputs of weather \citep{Crout20142}. The procedure for the SIMPLE model \citep{Zao2019simple} calibration  was a sequential process of modifying physiological variables specific for cocoa in the inputs files and  adding the appropriate files of weather for each region tested. In the SIMPLE model has seven input files where new cocoa crop data should be provided. Files in the list below can be edited to define the features of the new cultivars or experiments. Cocoa crop has not been simulated with SIMPLE model, hence consider previous cocoa studies, values such as leaf area index (LAI) \citep{Agele2016} and Harvest index (HI) \citep{Quintana2015} where modified. Radiation Use Efficiency (RUE) was calibrate according to the perennial crops  (banana and cotton) that had been calibrated previously in SIMPLE model \citep{Zao2019simple} with a RUE of 0.8 and 0.85 respectively. As cocoa trees are under shadow the RUE was lower with values between 0.7 and 0.5 g MJ$^{\mathbf{-1}}$ m$^{2}$ (table \ref{tab:reparam}).
Once the physiological parameters were calibrated to the  simulated yields for cocoa were reasonably close to the observed yield. The 23 flowering from Fedecacao reports were introduced in the treatment file, running the program and saving the results.

\subsection{Parameters}

{\color{olive}This cocoa model has  were three parameters which vary by region (table \ref{tab:reparam}): The thermal time required for harvest after the flowering date (Tsum) , the Radiation Use Efficiency (RUE) and yield observed on field. Physiological parameters in table \ref{tab:Treaparam} are  common  for all the regions studied. These parameters were calibrate for cultivars ICS95 and CCN51 considering a range of time of 200 DAF from flowering date to harvest day, even thought farmers collect the pod at 180 DAF. Heat and water stress parameters were not considered. }

@@

\subsection{Evaluation of Model Performance}

The cocoa model performance was evaluated by comparing simulated values cocoa yield with those reported by Fedecacao from cocoa plantations, using the statistical indice of relative root mean square error (RRMSE) (Equ. \ref{equ:RMSE}) \citep{Zao2019simple, Bai2020}.

@@

\section{RESULT}

\subsection{Weather Conditions Over Flowering Time }

{\color{olive} The weather data from NASA platform is showed in the figure \ref{fig:temp}. Data from 2018 to 2020 were analysed to show the tendencies by two years when the field data was reported.} Looking at the data, Santander and Caldas had the biggest variability and the maximum of solar radiation values over 20 MJ m$^{2}$day$^{-1}$. In contrast , Cali and Arauca presented the lowest values of PAR below 5 MJ m$^{2}$day$^{-1}$. (Fig.\ref{fig:temp}, a).  Even though, Cali and Santander had contrasting PAR conditions, they regions presented  have similar temperature during 2020 (Fig.\ref{fig:temp}, b). The temperature ranges from 16 to 28 $^\circ$C and it is relatively constant for each region. However, Arauca presented the biggest variability with hotter months during the first half of the year 2019 and 2020.  Apartado was found as the hottest region studied with 26 $^\circ$C and Caldas as the coldest site with 18 $^\circ$C.  Precipitation in Colombia is presented in two seasons per year from February to April and from October to November, while the relative humidity remain constant over 80\% (Fig.\ref{fig:temp},c). In general, the coldest regions tested (Caldas and Santander) had the maximun values of solar radiation available for photosynthesis.

\begin{figure}[h!]
	\centering
	\includegraphics[scale=0.4]{images/clima.png}
	\caption{\footnotesize {Colombian weather conditions. \\}}	
	\label{fig:temp}
	{\footnotesize (a), Available photosynthetic solar radiation (PAR). (b), Daily average temperature  (c), Monthly average of precipitation (bars) and relative humidity (dotted line) from 2018 to 2021 }
\end{figure}
\newpage

Figure \ref{fig:heat} shows the pearson correlation to study the weather data of the flowering time over the months of flowering  (monthF), month of harvest (monthH) and their final yield (fruit\_kg). The results showed that thermal time  T$_{b}$ (ttb) is 0.52 correlated with daily average temperature and  maximum temperature (TMAX)  and  temperature minimum (TMIN) and Dew Frost Point at 2 meters (T2MDEW) with  a correlation coefficient of 0.60. The wind (WS2M) is correlated with ttb with 0.57. However, less clear correlations were found of monthF  with T2MDEW, relative humidity (RH), WS2M and rain.

\begin{figure}[h!]
	\centering
	\includegraphics[scale=0.4]{images/heatm.png}\\
	\caption{\footnotesize {Pearson correlation average weather variables and flowering dates for five locations in Colombia. \\}}
	\label{fig:heat} 
	{\footnotesize Numbers in the squares are the correlation coefficients }
\end{figure}
\newpage

\subsection{Thermal Time }

Thermal time characterisation was made considering 180 DAF for each location. The boxplot in the figure \ref{fig:ttbox} shows the data distribution where boxes indicate the range of the central 50\% of the total data per region, the central line in the box is marking the median value and lines draw out from each box mean the range of the remaining data. Therefore, this boxplot shows differences between locations as was expected following the tendencies of the temperature  (fig. \ref{fig:temp},c). Apartado and Arauca had the highest temperatures hence, the highest thermal time values with 2909 and 2764 days$^\circ$C  respectively. Caldas had the lowest values with 1173 days$^\circ$C. Meanwhile, Cali and Santander presented similar thermal time around 2000 $^\circ$C (table \ref{tab:reparam}). The accumulated temperature during the pod development (fig.\ref{fig:ttbox}) depends on the region where cocoa is cultivated, and the variety planted. Thermal time values are also proportional to the yield reported on field (table \ref{tab:reparam}). \\
